---
title: "Data analysis"
author: "K.K."
date: "`r Sys.Date()`"
output:
  html_document:
    css: ../css/style.scss
    md_extensions: -ascii_identifiers
    theme: cerulean
    df_print: default
    chunk_output_type: console
    toc: yes
    toc_float:
      collapsed: false
---

<link rel="stylesheet" href="../js/highlight/styles/monokai.css">
<script src="../js/highlight/highlight.pack.js"></script>
<script>hljs.initHighlightingOnLoad();</script>

<style>
@import url('https://fonts.googleapis.com/css2?family=Montserrat:wght@400;700&display=swap');
</style>

# Packages

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotrix)
library(emmeans)
library(effectsize)
library(afex)
library(ggmosaic)
library(vcd)
source("../function/my_functions.R")
```

```{r, eval=FALSE, include=FALSE}
# Rearranging the original data
tmp <- read_csv("../data/matsumoto3.csv") %>%
  mutate(
    id = as.numeric(factor(number)),
    policy = if_else(kind_of_answer == "Opp", "risky", "sure"),
    outcome = case_when(
      outcome == "inequity" ~ "inequality",
      outcome == "best" ~ "success",
      outcome == "worst" ~ "failure",
      outcome == "unknown" ~ "control",
      TRUE ~ outcome
    ),
    outcome = if_else(policy == "sure", "sure", outcome),
    univ = if_else(univ == "t", "tokyo", "hokkaido"),
    order = if_else(Order == "Opp_First", "risky_sure", "sure_risky"),
    gender = if_else(sex == "m", "man", "woman")
  ) %>%
  rename(
    rating_policy = Eval,
    rating_distributor = Distribute,
    rating_recipient = Distributed
  ) %>%
  select(
    id, policy, outcome,
    rating_policy, rating_distributor, rating_recipient,
    order, univ, gender, age
  )

# Make the data frame containing each participant's information
tmp %>%
  select(
    id, rating_distributor, rating_recipient,
    order, univ, gender, age, outcome
  ) %>%
  distinct() %>%
  filter(outcome != "sure") %>%
  rename(cond = outcome) %>%
  mutate(type = case_when(
    rating_distributor >= 3 & rating_recipient >= 3 ~ "risky",
    rating_distributor <= 2 & rating_recipient <= 2 ~ "sure",
    TRUE ~ "other"
  )) %>%
  write_csv("../data/outcomebias_participants.csv")

tmp %>%
  select(id, policy, outcome, rating_policy) %>%
  write_csv("../data/outcomebias_task.csv")

rm(tmp)
```

# Participants

```{r, message=FALSE}
df_participants <- read_csv("../data/outcomebias_participants.csv")
```

## Demographic variables

```{r}
df_participants %>% count(univ, gender)
```

```{r}
df_participants %>%
  group_by(univ) %>%
  summarise(mean_age = mean(age), sd_age = sd(age), .groups = "drop") %>%
  mutate(mean_age = round(mean_age, 1), sd_age = round(sd_age, 1))
```

## Assignment to the conditions

```{r}
df_participants %>% count(cond)
```

# No significant difference in the policy evaluation between the universities

```{r, message=FALSE}
df_task <- read_csv("../data/outcomebias_task.csv")
```

## Figure 1b
```{r}
figure_1b <- df_task %>%
  left_join(df_participants, by = "id") %>%
  group_by(outcome, univ) %>%
  summarise(
    mean = mean(rating_policy), se = std.error(rating_policy),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = outcome, y = mean, color = univ, shape = univ)) +
  geom_point(size = 2, position = position_dodge(width = 0.4)) +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    width = 0.2, position = position_dodge(width = 0.4)
  ) +
  geom_vline(aes(xintercept = 4.5), linetype = "dashed") +
  scale_x_discrete(
    limits = c("success", "failure", "control", "inequality", "sure"),
    labels = c("Success", "Failure", "Control", "Inequality", "The sure\npolicy")
  ) +
  scale_y_continuous(breaks = seq(3, 4.5, 0.5), limits = c(3.4, 4.8)) +
  scale_color_manual(
    values = c("#23651c", "#f2a900"),
    labels = c("Hokkaido", "Tokyo")
  ) +
  scale_shape_discrete(labels = c("Hokkaido", "Tokyo")) +
  labs(
    x = "",
    y = "Ratings for the policies",
    color = "University",
    shape = "University"
  ) +
  theme(
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.95),
    legend.title = element_blank()
  )
figure_1b
```

## The risky policy

Two-way ANOVA (University * Outcome)

```{r}
# Two-way ANOVA (University * Outcome)
anova_risky_univ_outcome <- df_task %>%
  filter(outcome != "sure") %>%
  left_join(df_participants, by = "id") %>%
  aov(rating_policy ~ univ * outcome, data = .)

# ANOVA table
anova_risky_univ_outcome %>% summary()

# Effect sizes
anova_risky_univ_outcome %>% eta_squared()
```

## The sure policy

Welch's t-test

```{r}
# Welch's t-test
df_task %>%
  filter(outcome == "sure") %>%
  left_join(df_participants, by = "id") %>%
  t.test(rating_policy ~ univ, data = .)

# Effect size
df_task %>%
  filter(outcome == "sure") %>%
  left_join(df_participants, by = "id") %>%
  cohens_d(rating_policy ~ univ, data = .)
```

No significant differences were observed in the policy evaluation between the two universities. Thus, we merged data from the two universities.

# The unequal outcome biased the evaluation of the risky policy

## Figure 1c

```{r}
df_task %>%
  group_by(outcome) %>%
  summarise(
    mean = mean(rating_policy), se = std.error(rating_policy),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = outcome, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_vline(aes(xintercept = 4.5), linetype = "dashed") +
  scale_x_discrete(
    limits = c("success", "failure", "control", "inequality", "sure"),
    labels = c("Success", "Failure", "Control", "Inequality", "Sure")
  ) +
  scale_y_continuous(limits = c(3.5, 4.5), breaks = seq(3, 4.5, 0.5)) +
  labs(
    x = "Outcomes of the risky policy",
    y = "Ratings for the policies"
  )
```

## The risky policy

One-way ANOVA

```{r}
# One-way ANOVA (the four conditions)
anova_risky_outcome <- df_task %>%
  filter(outcome != "sure") %>%
  aov(rating_policy ~ outcome, data = .)

# ANOVA table
anova_risky_outcome %>% summary()

# Effect sizes
anova_risky_outcome %>% eta_squared()

# Multiple comparisons
anova_risky_outcome %>% emmeans(pairwise ~ outcome, adjust = "bon")
```

Effect sizes

```{r}
# Effect size for the multiple comparisons shown above
t_to_eta2(t = c(-3.427), df_error = 247)
```

## The risky vs. sure policy

Paired t-tests

```{r}
# Paired t-tests with Bonferroni correction
df_task_wider <- df_task %>%
  pivot_wider(
    id_cols = id,
    names_from = policy,
    values_from = c(outcome, rating_policy)
  ) %>%
  select(-outcome_sure) %>%
  rename(
    outcome = outcome_risky,
    rating_sure = rating_policy_sure,
    rating_risky = rating_policy_risky
  )

# Success
df_task_wider %>%
  filter(outcome == "success") %>%
  t.test(Pair(rating_sure, rating_risky) ~ 1, data = .)

# Effect size (success)
df_task_wider %>%
  filter(outcome == "success") %>%
  cohens_d(Pair(rating_sure, rating_risky) ~ 1, data = .)

# Failure
df_task_wider %>%
  filter(outcome == "failure") %>%
  t.test(Pair(rating_sure, rating_risky) ~ 1, data = .)

# Effect size (failure)
df_task_wider %>%
  filter(outcome == "failure") %>%
  cohens_d(Pair(rating_sure, rating_risky) ~ 1, data = .)

# Control
df_task_wider %>%
  filter(outcome == "control") %>%
  t.test(Pair(rating_sure, rating_risky) ~ 1, data = .)

# Effect size (control)
df_task_wider %>%
  filter(outcome == "control") %>%
  cohens_d(Pair(rating_sure, rating_risky) ~ 1, data = .)

# Inequality
df_task_wider %>%
  filter(outcome == "inequality") %>%
  t.test(Pair(rating_sure, rating_risky) ~ 1, data = .)

# Effect size (inequality)
df_task_wider %>%
  filter(outcome == "inequality") %>%
  cohens_d(Pair(rating_sure, rating_risky) ~ 1, data = .)

p.adjust(c(0.5192, 0.5776, 0.6701, 0.001135), method = "bonferroni")
```

Descriptive statistics

```{r}
# Mean
df_task_wider %>%
  group_by(outcome) %>%
  summarise(
    mean_risky = mean(rating_risky), mean_sure = mean(rating_sure),
    .groups = "drop"
  )
```

# Distributive preferences

## Figure 2a

```{r}
df_participants %>%
  mutate(
    distributor = if_else(rating_distributor >= 3, "opportunity", "outcome"),
    recipient = if_else(rating_recipient >= 3, "opportunity", "outcome")
  ) %>%
  count(distributor, recipient)

mosaic_plot <- df_participants %>%
  mutate(
    distributor = if_else(rating_distributor >= 3, "risky", "sure"),
    recipient = if_else(rating_recipient >= 3, "risky", "sure")
  ) %>%
  ggplot() +
  geom_mosaic(
    aes(x = product(distributor, recipient), fill = distributor),
    alpha = 1, show.legend = FALSE
  ) +
  labs(x = "Recipient", y = "Decision-maker")

mosaic_plot +
  geom_text(
    data = layer_data(mosaic_plot, 1) %>%
      select(xmin:ymax) %>%
      mutate(m.x = (xmin + xmax) / 2, m.y = (ymin + ymax) / 2) %>%
      select(m.x, m.y) %>%
      mutate(string = c("62", "32", "29", "128")),
    aes(x = m.x, y = m.y, label = string)
  )
```

Chi-square test and effect size

```{r}
chisq.test(matrix(c(62, 32, 29, 128), nrow = 2, byrow = TRUE))
cramers_v(matrix(c(62, 32, 29, 128), nrow = 2, byrow = TRUE))
```

Binomial tests

```{r}
binom.test(c(160, 91))
binom.test(c(157, 94))
```

## Figure 2b

```{r}
df_participants %>%
  filter(type != "other") %>%
  left_join(df_task, by = "id") %>%
  group_by(type, outcome) %>%
  summarise(
    mean = mean(rating_policy), se = std.error(rating_policy),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = outcome, y = mean, color = type, shape = type)) +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    position = position_dodge(width = 0.4), width = 0.2
  ) +
  geom_vline(aes(xintercept = 4.5), linetype = "dashed") +
  scale_x_discrete(
    limits = c("success", "failure", "control", "inequality", "sure"),
    labels = c("Success", "Failure", "Control", "Inequality", "Sure")
  ) +
  labs(
    x = "Outcomes of the risky policy",
    y = "Ratings for the policies"
  )
```

## The risky policy

Two-way ANOVA

```{r}
# Two-way ANOVA
anova_risky_preference <- df_participants %>%
  mutate(
    distributor = if_else(rating_distributor >= 3, "risky", "sure"),
    recipient = if_else(rating_recipient >= 3, "risky", "sure"),
  ) %>%
  left_join(df_task, by = "id") %>%
  filter(distributor == recipient) %>%
  filter(outcome != "sure") %>%
  aov(rating_policy ~ outcome * distributor, data = .)

# ANOVA table
anova_risky_preference %>% summary()

anova_risky_preference %>% eta_squared()

anova_risky_preference %>% 
  emmeans(specs = pairwise ~ distributor * outcome)

```

```{r}
t_to_eta2(
  t = c(3.895, -2.944, 0.457), df_error = 182
)
```

## The sure policy

```{r}
df_participants %>%
  mutate(
    distributor = if_else(rating_distributor >= 3, "risky", "sure"),
    recipient = if_else(rating_recipient >= 3, "risky", "sure"),
  ) %>%
  left_join(df_task, by = "id") %>%
  filter(distributor == recipient) %>%
  filter(outcome == "sure") %>%
  t.test(rating_policy ~ distributor, data = .)

df_participants %>%
  mutate(
    distributor = if_else(rating_distributor >= 3, "risky", "sure"),
    recipient = if_else(rating_recipient >= 3, "risky", "sure"),
  ) %>%
  left_join(df_task, by = "id") %>%
  filter(distributor == recipient) %>%
  filter(outcome == "sure") %>%
  cohens_d(rating_policy ~ distributor, data = .)
```